# Längsschnittliche Messinvarianz

## Konfigurale Invarianz
```{r}
library(lavaan)
library(semTools)
library(semPlot)
library(haven)

# read complete data
data_star_selected_variables <- read_spss("data/data_star_selected_variables.sav")

part_configinv_mod <- 
  "part4e =~ g4ptattn + g4ptmtrl + g4ptpers + g4ptlate
   part8e =~ g8peattn + g8pemtrl + g8pepers + g8pelate

   # latent (co-)variances
   part4e ~~ part4e
   part8e ~~ part8e
   part4e ~~ part8e
   
   # residual variances
   g4ptattn ~~ g4ptattn
   g4ptmtrl ~~ g4ptmtrl
   g4ptpers ~~ g4ptpers
   g4ptlate ~~ g4ptlate
   g8peattn ~~ g8peattn
   g8pemtrl ~~ g8pemtrl
   g8pepers ~~ g8pepers
   g8pelate ~~ g8pelate
   
   # set residual covariances free over time per indicator
   g4ptattn ~~ g8peattn
   g4ptmtrl ~~ g8pemtrl
   g4ptpers ~~ g8pepers
   g4ptlate ~~ g8pelate
   "

part_configinv_fit <- 
    cfa(part_configinv_mod, data = data_star_selected_variables)

semPaths(part_configinv_fit)
```

## Schwache Invarianz
```{r}
part_weakinv_mod <- 
  "part4e =~ g4ptattn + s1*g4ptmtrl + s2*g4ptpers + s3*g4ptlate 
   part8e =~ g8peattn + s1*g8pemtrl + s2*g8pepers + s3*g8pelate

   # latent (co-)variances
   part4e ~~ part4e
   part8e ~~ part8e
   part4e ~~ part8e
   
   # residual variances
   g4ptattn ~~ g4ptattn
   g4ptmtrl ~~ g4ptmtrl
   g4ptpers ~~ g4ptpers
   g4ptlate ~~ g4ptlate
   g8peattn ~~ g8peattn
   g8pemtrl ~~ g8pemtrl
   g8pepers ~~ g8pepers
   g8pelate ~~ g8pelate
   
   # set residual covariances free over time per indicator
   g4ptattn ~~ g8peattn
   g4ptmtrl ~~ g8pemtrl
   g4ptpers ~~ g8pepers
   g4ptlate ~~ g8pelate
   "

part_weakinv_fit <- 
    cfa(part_weakinv_mod, data = data_star_selected_variables)

semPaths(part_weakinv_fit, what = "col")
semPaths(part_weakinv_fit, what = "est", fade = F)

summary(compareFit(part_configinv_fit, part_weakinv_fit))
```



## Starke Invarianz
```{r}
part_stronginv_mod <- 
  "part4e =~ g4ptattn + s1*g4ptmtrl + s2*g4ptpers + s3*g4ptlate 
   part8e =~ g8peattn + s1*g8pemtrl + s2*g8pepers + s3*g8pelate

   # latent (co-)variances with constrains
   part4e ~~ v1*part4e
   part8e ~~ v1*part8e
   part4e ~~ part8e
   
   # residual variances
   g4ptattn ~~ g4ptattn
   g4ptmtrl ~~ g4ptmtrl
   g4ptpers ~~ g4ptpers
   g4ptlate ~~ g4ptlate
   g8peattn ~~ g8peattn
   g8pemtrl ~~ g8pemtrl
   g8pepers ~~ g8pepers
   g8pelate ~~ g8pelate
   
   # set residual covariances free over time per indicator
   g4ptattn ~~ g8peattn
   g4ptmtrl ~~ g8pemtrl
   g4ptpers ~~ g8pepers
   g4ptlate ~~ g8pelate
   "

part_stronginv_fit <- 
    cfa(part_stronginv_mod, data = data_star_selected_variables)

semPaths(part_stronginv_fit, what = "col")
semPaths(part_stronginv_fit, what = "est", fade = F)

summary(compareFit(part_configinv_fit, part_weakinv_fit, part_stronginv_fit))
```


## Strikte Invarianz
```{r}
part_strictinv_mod <- 
  "part4e =~ g4ptattn + s1*g4ptmtrl + s2*g4ptpers + s3*g4ptlate 
   part8e =~ g8peattn + s1*g8pemtrl + s2*g8pepers + s3*g8pelate

   # latent (co-)variances with constrains
   part4e ~~ v1*part4e
   part8e ~~ v1*part8e
   part4e ~~ part8e
   
   # residual variances
   g4ptattn ~~ r1*g4ptattn
   g4ptmtrl ~~ r2*g4ptmtrl
   g4ptpers ~~ r3*g4ptpers
   g4ptlate ~~ r4*g4ptlate
   g8peattn ~~ r1*g8peattn
   g8pemtrl ~~ r2*g8pemtrl
   g8pepers ~~ r3*g8pepers
   g8pelate ~~ r4*g8pelate
   
   # set residual covariances free over time per indicator
   g4ptattn ~~ g8peattn
   g4ptmtrl ~~ g8pemtrl
   g4ptpers ~~ g8pepers
   g4ptlate ~~ g8pelate
   "

part_strictinv_fit <- 
    cfa(part_strictinv_mod, data = data_star_selected_variables)

semPaths(part_strictinv_fit, what = "col")
semPaths(part_strictinv_fit, what = "est", fade = F)

summary(compareFit(part_configinv_fit, part_weakinv_fit, 
                   part_stronginv_fit, part_strictinv_fit))
```


## Weiterführende Literatur
:::{.callout-note collapse=false appearance='default' icon=true}
## Literaturempfehlungen zum Thema Longitudinal Measurement Invariance
* Mackinnon, S., Curtis, R., & O’Connor, R. (2024). A Tutorial in Longitudinal Measurement Invariance and Cross-lagged Panel Models Using lavaan. https://doi.org/10.31234/osf.io/tkzrb
* Brown, T. A. (2015). Confirmatory factor analysis for applied research. Guilford Press.
:::